---
title: "Aflevering mandag 23. september"
author: "Niklas Jespersen"
date: "21/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(colorspace)
library(tidyverse)
library(dplyr)
library(httr)
library(rjstat)
library(conflicted)
library(ggplot2)
library(zoo)

```


```{r}
conflict_prefer("filter", "dplyr")
```


Saving the links as to varibles, link76 and link74

```{r}
link76="http://data.ssb.no/api/v0/dataset/95276.json?lang=no"
link74="http://data.ssb.no/api/v0/dataset/95274.json?lang=no"

```



```{r}
d.temp = GET(link76)

Tabel1 = fromJSONstat(content(d.temp, "text"))

Tabl_76= Tabel1[[1]]
```


```{r}
d.temp2 = GET(link74)

Tabel2 = fromJSONstat(content(d.temp2, "text"))

Tabl_74= Tabel2[[1]]
```




Merging the two dataframes together with merge

```{r}
Data_hotels = rbind(Tabl_74, Tabl_76)

```

Renaming all the varible names to english
Renaming all row names

```{r}
Data_hotels = rename(Data_hotels, "Region"="region")

Data_hotels = rename(Data_hotels, "Date"="måned")

Data_hotels = rename(Data_hotels, "Variable"="statistikkvariabel")

Data_hotels = rename(Data_hotels, "Value"="value")



Data_hotels$Date = as.Date(as.yearmon(Data_hotels$Date, "%YM%m"))



Data_hotels$Variable <- recode(Data_hotels$Variable, "Kapasitetsutnytting av senger (prosent)" = "Beds percentage capacity utilization", "Kapasitetsutnytting av rom (prosent)" = 	

"Rooms percentage capacity utilization", "Pris per rom (kr)" = "Price per room")
```

Filtering out data with the price 

```{r}
Price_DATA=Data_hotels %>% filter(Value != 0) %>% filter(Variable=="Price per room")

```


I caluculate the Region average

```{r}
Region_average = Price_DATA %>% group_by(Region) %>% summarize(average_price=mean(Value))
```



I calculate the averege of "Hele landet"

```{r}
hele_landet = Price_DATA %>% filter(Region == "Hele landet")
```

```{r}
Mean_hele_landet = mean(hele_landet$Value)
```


I find the difference between the county average room price and the national average room price per month


```{r}
Dif_meanRoomPrice = Price_DATA %>% group_by(Region) %>% summarize(average_price=mean(Value)-Mean_hele_landet)

```


I find find the amount the region with the higest average price compared to the national average.

```{r}
L = Dif_meanRoomPrice$average_price == max(Dif_meanRoomPrice$average_price)

paste( "The higest average price compared to the national average is", Dif_meanRoomPrice[L,]$average_price, "and the region is", Dif_meanRoomPrice[L,]$Region)

```


I find find the amount the region with the lowest average price compared to the national average.

```{r}
M <- Dif_meanRoomPrice$average_price == min(Dif_meanRoomPrice$average_price)

paste( "The lowest average price compared to the national average is", Dif_meanRoomPrice[M,]$average_price, "and the region is", Dif_meanRoomPrice[M,]$Region)
```


##Do the calculations above per county and per year. Identify the year and county with the highest positive and negative difference.


```{r}
Average_Price_By_Year_Region= Price_DATA %>% group_by(lubridate::year(Date), Region) %>% summarize(average_price_year=mean(Value))


p = Average_Price_By_Year_Region$average_price_year == max(Average_Price_By_Year_Region$average_price_year)

paste( "The higest average price is", Average_Price_By_Year_Region[p,]$average_price_year, "and the region is", Average_Price_By_Year_Region[p,]$Region, "in year", Average_Price_By_Year_Region[p,]$`lubridate::year(Date)` )

d = Average_Price_By_Year_Region$average_price_year == min(Average_Price_By_Year_Region$average_price_year)

paste( "The lowest average price is", Average_Price_By_Year_Region[d,]$average_price_year, "and the region is", Average_Price_By_Year_Region[d,]$Region, "in year", Average_Price_By_Year_Region[d,]$`lubridate::year(Date)` )


```

Make a plot of the monthly price difference for Troms county from 1999 until today.


```{r}
Plot_Data= Price_DATA %>% filter(Region=="Troms - Romsa", Date >= "1999-01-01" & Date <= "2019-07-01")

ggplot(Plot_Data) +
  geom_point(aes(x = Date, y = Value)) +
  labs(
    x = "Date of the year",
    y = "Price in Troms",
    title = "Over time the price in Troms has increased",
    subtitle = "Prices in Troms from 1999 to 2019",
    caption = "Assignment 2")
```


Per county, is there any relationship (correlation) between room capacity and price since January 2010?

I tidy up my data

```{r}
Datahotel1=Data_hotels %>% filter(Value != 0)

Clean_data=na.omit(spread(Datahotel1, Variable, Value, fill = NA, convert = FALSE))

```

I find the correlations per Region

```{r}
Clean_data %>%
  filter(Date >= "2010-01-01" & Date <= "2019-07-01") %>%
  group_by(Region) %>%
  summarize(COR=cor(`Price per room`, `Rooms percentage capacity utilization`))
```






